Stage-Splitting a Modal Language

### 2014-10-30 paper/splitting.tex
small edits

Splitting a stage-\bbtwo\ term $\coltwo e A$ produces a \emph{precomputation}
term $p$ and a \emph{resumer} function $l.r$, such that $e$ is equivalent to
$\letin l p r$. This is implemented by the judgment $\splittwosym$,
(\cref{fig:termSplitTwo}), which operates inductively on the structure of the
term.

@@

Since base types and variables have no stage \bbone\ content, their
precomputations are \texttt{()}, and are not used in the resumer. At every other
expression except $\prev$, splitting bundles the precomputations of the
constituent parts, and then unbundles them with a pattern, with the resumer
precisely resembling the original expression. The $\prev$ rule depends on
splitting stage \bbone\ terms, which is covered later.

@@

The same is not true of partial values (at stage \bbone), which \emph{can}
contain stage \bbtwo\ computations, namely, inside $\next$ blocks, which must
contain a variable reference to some future context, and inside lambdas, which
may contain a $\next$ in their body.

Thus, to split partial values, we must separate the contents of $\next$ blocks
from the rest of the term, while maintaining the term's original structure. This
is implemented by the judgment $v \vsplito [v_1,v_2]$ (\ref{fig:valSplit}).
At product, sum, and recursive types, splitting acts recursively, gathering all
subterms' stage $w$ content into $v_w$.
At base types, splitting assigns $v$ entirely to stage \bbone\ (and \texttt{()}
to stage \bbtwo);
on the other hand, at $\fut$ types, splitting assigns $v$ (which must have the
form $\next~\hat y$) entirely to stage \bbtwo\ (and \texttt{()} to stage
\bbone).
This process of exposing only the relevant structure of a term at each stage is
called \emph{masking}.

@@

@@

\paragraph {Speculation}

In the stage \bbtwo\ rules for {\tt if}s, {\tt case}s, and functions, the
precomputation is lifted out from within branches. This is the manifestation of
the speculation behavior from the semantics.

>>>>

Splitting a stage \bbtwo\ term $\coltwo e A$ produces a \emph{precomputation}
term $p$ and a \emph{resumer} function $l.r$, such that $e$ is equivalent to
$\letin l p r$. This is implemented by the judgment $\splittwosym$ 
(\cref{fig:termSplitTwo}), which operates inductively on the structure of the
term.

@@

Since base types and variables have no stage \bbone\ content, their
precomputations are \texttt{()}, and are not used in the resumer. At every other
expression except $\prev$ and $\pause$, splitting bundles the precomputations of the
constituent parts, and then unbundles them with a pattern, with the resumer
precisely resembling the original expression. The $\prev$ and $\pause$ rules depends on
splitting stage \bbone\ terms, which is covered later.

Observe that in the rules for {\tt if}s, {\tt case}s, and functions, the
precomputation is lifted out from within branches. This is the manifestation of
the speculation behavior from the semantics.

@@

The same is not true of partial values (by definition, stage \bbone), which \emph{can}
contain stage \bbtwo\ computations, namely, inside $\next$ blocks, which must
contain a variable reference to some future context, and inside lambdas, which
may contain a $\next$ in their body.

Thus, to split partial values, we must separate the contents of $\next$ blocks
from the rest of the term, while maintaining the term's original structure. This
is implemented by the judgment $v \vsplito [v_1,v_2]$ (\ref{fig:valSplit}).
At product, sum, and recursive types, splitting acts recursively, gathering all
subterms' stage $w$ content into $v_w$.
At base types, splitting assigns $v$ entirely to stage \bbone\ (and \texttt{()}
to stage \bbtwo);
conversely, at $\fut$ types, splitting assigns $v$ (which must have the
form $\next~\hat y$) entirely to stage \bbtwo\ (and \texttt{()} to stage
\bbone).
This process of exposing only the relevant structure of a term at each stage is
called \emph{masking}.

@@

\TODO talk about the changed function signature, which is necessitated by the precomputation;
note that we have to assume {\em all} stage \bbone\ functions have meaningful precomputations

@@
